---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

```{r}
library(readxl)
library(tidyverse)
library(car)

Concrete_Data <- read_excel("C:/Users/HP DV6/Downloads/Concrete_Data.xls")

# Rename column name to make it easier to operate on.
Concrete_Data <- rename(Concrete_Data,
  concrete = `Concrete compressive strength(MPa, megapascals)`,
  cement = `Cement (component 1)(kg in a m^3 mixture)`,
  b_furnace = `Blast Furnace Slag (component 2)(kg in a m^3 mixture)`,
  water = `Water  (component 4)(kg in a m^3 mixture)`,
  superpl = `Superplasticizer (component 5)(kg in a m^3 mixture)`,
  c_aggregate = `Coarse Aggregate  (component 6)(kg in a m^3 mixture)`,
  f_aggregate = `Fine Aggregate (component 7)(kg in a m^3 mixture)`,
  age = `Age (day)`,
  fly_ash = `Fly Ash (component 3)(kg in a m^3 mixture)`
  
  
)
train <- Concrete_Data[1:900,]
test <- Concrete_Data[901:1030,]

```

Graphical analysis of concrete data set.
```{r}

```

Examining bivariate relationship between dependent variables to determine if interaction effect exist.

```{r}
cor(Concrete_Data)
scatterplotMatrix(Concrete_Data, spread=FALSE, smoother.args=list(lty=2),
main="Scatter Plot Matrix")

```

Fitting a muitlple linear regression on the data set.
```{r}
lm_fit <- lm(concrete ~ ., data = train)
summary(lm_fit)
```

There is a relationship between the response and predicator variables with a F-statistic of 176.2 and a p-value of < 2.2e-16.
Most of the predicator variables are statistically significant save two, c_aggregate and f_aggregate.Overall, this model accounts for 60.92% of concrete's variance.

Diagnosing the model.
```{r}
plot(lm_fit)
```
The Residuals vs Fitted plot shows elements of heteroscedasticity. The plot also shows 382, 384 and 225 are outliers.
The Residuals vs Leverage shows 225 ,711, and 55 have high leverages.

Removing outliers, variables with high leverage, and also log transforming the response variable.
```{r}
train <- filter(train, !cement %in% c(315,516,212.52,305.3,189.6))
lm_fit2 <- lm(concrete~.-c_aggregate-f_aggregate, train)
summary(lm_fit2)
plot(lm_fit2)
```
After removing outliers the model R-squared increased from 0.6092 to 0.6129, and the RSE reduces from 10.76 to 10.7. 

Fitting interactions into the model.
```{r}
lm_fit3 <- lm(concrete~.-c_aggregate -f_aggregate + superpl : water    , train)
summary(lm_fit3)
plot(lm_fit3)
```
After fitting the interaction effect into the dataset. The model accounts for 62.36 % of the variance in concrete compressive strength.

Testing the model and compute the R^2 of the predicted values.
```{r}

predicted_values <- predict(lm_fit3, test)
predicted_values <- data.frame(predicted_values)
predicted_values <- cbind(predicted_values, test$concrete)

SSE <- sum((predicted_values$`test$concrete` - predicted_values$predicted_values  ) ^ 2)
SST  <- sum((predicted_values$`test$concrete` - mean(predicted_values$`test$concrete`)) ^ 2)


print(1 - SSE/ SST)

```
The linear model explained 56.40% of test data's concrete varaince.
